---
title: "Stats 306, Winter 2024, Final Group Project"
author: "Kaitlyn Anderson + 2 group members"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(maps)
library(reshape2)
library(modelr)
```


This data is downloaded from: <https://www.iea.org/data-and-statistics/data-tools/global-ev-data-explorer>


# Data Cleaning


```{r}

clean_df1 <- df1[, !(names(df1) %in% c("mode", "category", "unit"))]
clean_df2 <- df2[, !(names(df2) %in% c("mode", "category", "unit"))]

```


# Data Tidying


```{r}

wider_clean_df1 <- clean_df1 |> pivot_wider(names_from = 
                                              c(parameter, powertrain), 
                                            values_from = value)
wider_clean_df2 <- clean_df2 |> pivot_wider(names_from = 
                                              c(parameter, powertrain),
                                            values_from = value)

```


# Recreating visualizations from this article: <https://ourworldindata.org/electric-car-sales#article-licence>


# Diagram 01 and 02
```{r}

wider_clean_df1 |> 
  filter(region %in% 
           c("World", "Norway", "United Kingdom", "EU27", "China", "USA")) |> 
  ggplot(aes(x = year, y = `EV sales share_EV`)) +
  geom_line(aes(color = "red")) +
  labs(title = "Share of new cars sold that are electric, 2010 to 2022",
       x = "Year",
       y = "EV Sales Share (%)") +
  theme_minimal() +
  facet_wrap(~region, scales = "free_x", labeller = labeller(region = c(
               "World" = "World",
               "Norway" = "Norway",
               "United Kingdom" = "United Kingdom",
               "EU27" = "European Union (27)",
               "China" = "China",
               "USA" = "United States"))) + 
  ylim(0, 90) + scale_x_continuous(breaks = seq(2010, 2022, by = 2)) + 
  theme(axis.text.x = element_text(size = 6))

world_map <- map_data("world")
world_map <- world_map |> filter(region != "Antarctica")

data_2022 <- wider_clean_df1 |> filter(year == 2022)

renamed_data_2022 <- data_2022 |> 
  mutate(region = ifelse(region == "United Kingdom", "UK", region))

merged_data <- merge(world_map, renamed_data_2022, by.x = "region", 
                     by.y = "region", all.x = TRUE)

merged_data <- merged_data |>
  arrange(region, order)

ggplot(merged_data, aes(x = long, y = lat, group = interaction(region, group))) +
  geom_polygon(aes(fill = `EV sales share_EV`), color = NA, size = 0.1) + 
  geom_path(color = "black", size = 0.1) +
  scale_fill_gradient(low = "lightblue1", high = "darkgreen", 
                      na.value = "grey90", name = "EV Sales Share (%)") +
  labs(title = "Share of new cars sold that are electric, 2022") +
  theme_void() + 
  coord_fixed() + theme(legend.position = "bottom") +
  guides(fill = guide_colourbar(title.position = "top", 
                                title.hjust = 0.5, 
                                barwidth = 20, 
                                barheight = 0.5, 
                                horizontal = TRUE)) 


```

These line graphs show the percentage of vehicle sales that are electric vehicles sales 
from 2010 to 2022 in China, the EU, Norway, the UK, the US, and the World, and the map 
shows the electric vehicle sales share across all the regions included in the dataset 
in 2022. Out of all the regions, it is evident that Norway has had the greatest growth 
in EV sales shares from 2010 to 2022 and has the greatest percentage of EV sales in 2022 
compared to the other regions. Furthermore, Northern Europe and China have the greatest 
share of new cars that are electric in 2022. However, the other regions are also 
experiencing growth in EV sales shares, just at a slower rate than Northern Europe and China. 
# Diagram 03 and 04
```{r}

wider_clean_df1 |> 
  filter(region %in% 
           c("World", "Norway", "United Kingdom", "EU27", "China", 
             "USA", "Sweden", "Germany", "India", "South Africa")) |> 
  filter(year == 2022) |> mutate(region = reorder(region, `EV sales share_EV`)) |>
  ggplot(aes(x = region, y = `EV sales share_EV`)) +
  geom_col(fill = "skyblue3") +
  labs(title = "Share of new cars sold that are electric, 2022",
       x = "Region",
       y = "EV Sales Share (%)") +
  theme_minimal() + coord_flip()

world_map <- map_data("world")
world_map <- world_map |> filter(region != "Antarctica")

merged_data <- merge(world_map, renamed_data_2022, by.x = "region", 
                     by.y = "region", all.x = TRUE)

merged_data <- merged_data |>
  arrange(region, order)

ggplot(merged_data, aes(x = long, y = lat, group = interaction(region, group))) +
  geom_polygon(aes(fill = `EV sales share_EV`), color = NA, size = 0.1) + 
  geom_path(color = "black", size = 0.1) +
  scale_fill_gradient(low = "lightblue1", high = "darkgreen", 
                      na.value = "grey90", name = "EV Sales Share (%)") +
  labs(title = "Share of new cars sold that are electric, 2022") +
  theme_void() + 
  coord_fixed() + theme(legend.position = "bottom") +
  guides(fill = guide_colourbar(title.position = "top", 
                                title.hjust = 0.5, 
                                barwidth = 20, 
                                barheight = 0.5, 
                                horizontal = TRUE)) 

```

This bar chart shows the percentage of new cars sold that are electric in Norwary, Sweden, 
Germany, China, the UK, the EU, the World, the US, India, and South Africa in 2022. The 
x-axis of the graph is rearranged so that the regions are ordered by EV sales share. This
makes it clear to see that out of these 10 regions, Norway, Sweden, and Germany have the 
largest EV sales share, while the US, India, and South Africa have the lowest. China, the
UK, the EU, and the World moderate share of electric vehicle sales. The map is the same 
map as seen above, showing the electric vehicle sales share across all the regions included 
in the dataset in 2022.

# Diagram 05
```{r}

sales_data_processed <- wider_clean_df1 |>
  filter(region %in% c("World", "Norway", "United Kingdom", "EU27", "China", "USA")) |>
  mutate(EV_total_sales = (`EV sales_BEV` + `EV sales_PHEV`) / (`EV sales share_EV` / 100),
         proportion_BEV = (`EV sales_BEV` / EV_total_sales) * 100,
         proportion_PHEV = (`EV sales_PHEV` / EV_total_sales) * 100) |>
  select(region, year, proportion_BEV, proportion_PHEV)

long_data <- sales_data_processed |>
  pivot_longer(cols = c(proportion_BEV, proportion_PHEV), 
               names_to = "vehicle_type", 
               values_to = "sales_percentage")

ggplot(long_data, aes(x = year, y = sales_percentage, fill = vehicle_type)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("proportion_BEV" = "indianred3", 
                               "proportion_PHEV" = "skyblue3"),
                    labels = c("Battery-electric", "Plug-in hybrid"),
                    name = "Vehicle Type") +
  labs(title = 
         "Percentage of new cars sold that are battery-electric and plug-in hybrid, 2010 to 2022",
       x = "Year",
       y = "Percentage of total car sales") +
  facet_wrap(~ region, ncol = 3, scales = "free_x") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = seq(2010, 2022, by = 2)) +
  theme(axis.text.x = element_text(size = 6)) + theme(title = element_text(size = 9))

```
 
This stacked bar chart shows the percentage of new cars that are battery-electric and 
plug-in hybrid from 2010 to 2022 in China, the EU, Norway, the UK, the US, and the 
World. This graph illustrates that, overall, the percentage of new cars sold that are 
electric vehicles is growing over time. Furthermore, it shows that, of electric vehicle 
sales, battery-electric vehicles are more popular than plug-in hybrid vehicles in these 
regions. The percentage of  battery-electric vehicle sales seems to be growing at a 
faster rate than the percentage of plug-in hybrid vehicle sales, and more battery-electric 
vehicles are being sold than plug-in hybrid vehicles.

# Diagram 06 and 07
```{r}

wider_clean_df1 <- wider_clean_df1 |>
  mutate(EV_total_cars = `EV sales_BEV` + `EV sales_PHEV`) |>
  mutate(percentage_BEV = `EV sales_BEV` / EV_total_cars)

fourth_graph_data <- wider_clean_df1 |>
  filter(region %in% c("World", "Norway", "United Kingdom", "China", "Sweden")) |>
  filter(year >= 2012)

ggplot(fourth_graph_data, aes(x = year, y = percentage_BEV, color = region)) +
  geom_line() +
  geom_point() +
  labs(title = "Share of new electric cars that are fully battery-electric, 2012 to 2022",
       x = "Year",
       y = "Percentage of EV Sales that are BEV") +
  scale_color_manual(values = c("World" = "hotpink", "Norway" = "darkgreen", 
                                "United Kingdom" = "purple", 
                                "China" = "red", "Sweden" = "orange")) +
  scale_x_continuous(breaks = seq(2012, max(fourth_graph_data$year), by =
                                  2)) +
   scale_y_continuous(labels = c("0%", "20%", "40%", "60%", "80%",  
                      "100%"), breaks = seq(0, 1, by = 0.2)) +
  theme_minimal()

BEV_data_2022 <- wider_clean_df1 |> filter(year == 2022)

merged_BEV_data <- merge(world_map, wider_clean_df1, by.x = "region", 
                     by.y = "region", all.x = TRUE)

merged_BEV_data <- merged_BEV_data |>
  arrange(region, order)

ggplot(merged_BEV_data, aes(x = long, y = lat, group = interaction(region, group))) +
  geom_polygon(aes(fill = percentage_BEV), color = NA, size = 0.1) + 
  geom_path(color = "black", size = 0.1) +
  scale_fill_gradient(low = "lightyellow", high = "darkgreen", 
                      na.value = "grey90", name = "EV Sales Share (%)") +
  labs(title = "Share of new electric cars that are fully battery-electric, 2012 to 2022") +
  theme_void() + 
  coord_fixed() + theme(legend.position = "bottom") +
  guides(fill = guide_colourbar(title.position = "top", 
                                title.hjust = 0.5, 
                                barwidth = 20, 
                                barheight = 0.5, 
                                horizontal = TRUE))


```
 
This line graph depicts the share of new electric cars that are fully battery-electric 
from 2012 to 2022. There seemed to have been a slight trend around 2017 where more 
hybrids were bought, but overall, the percentage that are fully electric has been 
increasing. Sweden is the most interesting insight from this graph, since they are high 
on the list for electric car sales, but they are majority hybrids. 

# Diagram 08 and 09
```{r}

fifth_graph_regions <- wider_clean_df1 |>
  filter(region %in% c("World", "China", "EU27", "USA"))
  
  ggplot(fifth_graph_regions, aes(x = year, y = EV_total_cars, color = region)) +
  geom_line() +
  geom_point() +
  labs(title = "Number of new electric cars sold, 2010 to 2022",
       x = "Year",
       y = "Total Electric Cars") +
  scale_color_manual(values = c("World" = "orange", 
                                "EU27" = "darkgreen", "China" = "red", "USA" = "purple")) +
  scale_x_continuous(breaks = seq(2010, max(wider_clean_df1$year), by = 2)) +
  scale_y_continuous(labels = c("0", "2 million", "4 million", 
                                "6 million", "8 million", "10 million"), 
                      breaks = seq(0, 10 * 10^6, by = 2 * 10^6)) +
  theme_minimal()

merged_EV_data <- merge(world_map, BEV_data_2022, by.x = "region", 
                     by.y = "region", all.x = TRUE)

merged_EV_data <- merged_EV_data |>
  arrange(region, order)

ggplot(merged_EV_data, aes(x = long, y = lat, group = interaction(region, group))) +
  geom_polygon(aes(fill = EV_total_cars), color = NA, size = 0.1) + 
  geom_path(color = "black", size = 0.1) +
  scale_fill_gradient(low = "bisque", high = "darkred", na.value = "grey90", 
                      name = "EV Total Cars", labels = scales::comma) +  
  labs(title = "Number of new electric cars sold, 2010 to 2022") +
  theme_void() + 
  coord_fixed() + 
  theme(legend.position = "bottom") +
  guides(fill = guide_colourbar(title.position = "top", 
                                title.hjust = 0.5, 
                                barwidth = 20, 
                                barheight = 0.5, 
                                horizontal = TRUE))

```
 
This line graph shows the number of new electric cars sold, including both fully electric 
and hybrid. For all regions shown in the graph, the number has been increasing, particularly 
rapidly from 2020 on. China has shown the most individual increase in the number of sales, 
with the only region above it being the whole world. 

# Diagram 10
```{r}

df1_world_totals <- wider_clean_df1 |> 
  filter(region %in% "World") |>
  mutate(EV_total_sales = `EV sales_BEV` + `EV sales_PHEV`) |>
  mutate(total_sales = EV_total_sales / `EV sales share_EV` * 100) |>
  mutate(non_EV_total_sales = total_sales - EV_total_sales)

long_world_totals <- df1_world_totals |>
  pivot_longer(cols = c(EV_total_sales, non_EV_total_sales), 
               names_to = "vehicle_type", 
               values_to = "sale_numbers")

ggplot(long_world_totals, aes(x = year, y = sale_numbers, fill = vehicle_type)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("darkorange", "grey"),
                    labels = c("Electric cars", "Non-electric cars"),
                    name = "Vehicle Type") +
  labs(title = "Number of new cars sold, by type, World",
       x = "Year",
       y = "Total Cars Sold") +
  scale_y_continuous(labels = c("0 cars", "20 million cars", "40 million cars", 
                                "60 million cars", "80 million cars"),
                     breaks = seq(0, 80 * 10^6, by = 20 * 10^6)) +
  scale_x_continuous(breaks = unique(long_world_totals$year)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```
 
This bar graph shows the number of new cars sold in the world for each year broken down by 
either non-electric or electric. This graph shows that while the large majority of new cars
sold are non-electric, the number of non-electric cars sold peaked in 2018 and the proportion
of new cars sold that are electric has been increasing over time. 

# Diagram 11 and 12
```{r}

wider_clean_df1 |> filter(region %in% 
                            c('Norway', 'Sweden', 'China', 'World', 'USA')) -> p7
ggplot(p7, aes(x=year, y=`EV stock share_EV` / 100, group = region, color = region)) +
  geom_line() + geom_point() + scale_x_continuous(breaks = seq(2010, 2022)) +
  scale_y_continuous(
    labels = scales::percent_format(scale = 100), 
    limits = c(0, 0.3), 
    breaks = seq(0, 0.3, by = 0.05)) +
  labs(
    title = "Share of cars currently in use that are electric, 2010 to 2022",
    x = "Year",
    y = "EV Stock Share",
    color = "Region")

world_map <- map_data("world") |> filter(region != 'Antarctica')
year_2022_data = wider_clean_df1 |> filter(year == 2022)
merged_data = merge(world_map, year_2022_data, by.x = 'region', 
                    by.y = 'region', all.x = T)
merged_data <- merged_data |>
  arrange(region, order)

ggplot(merged_data, aes(x = long, y = lat, group = interaction(region, group))) +
  geom_polygon(aes(fill = `EV stock share_EV`), color = NA, size = 0.1) + 
  geom_path(color = "black", linewidth = 0.1) +
  scale_fill_gradient(low = "lightblue1", high = "darkgreen", 
                      na.value = "grey90", name = "EV Stock Share (%)") +
  labs(title = "Share of cars currently in use that are electric, 2010 to 2022") +
  theme_void() + 
  coord_fixed() + theme(legend.position = "bottom") +
  guides(fill = guide_colourbar(title.position = "top", 
                                title.hjust = 0.5, 
                                barwidth = 20, 
                                barheight = 0.5, 
                                horizontal = TRUE))

```
 
This scatterplot and line graph depicts the percentage share of electric vehicles in 
use out of the total car stock for different regions from 2010 to 2022. Out of all 
plotted regions, Norway seems to have the highest percentage of electric vehicle stock
in use, with the share increasing to over 25% by 2022. Other regions follow similar
trends with the exception of doing so at a more gradual pace. 

# Diagram 13 and 14
```{r}

wider_clean_df1 |> filter(region %in% c('World', 'China', 'EU27', 'USA')) -> p8
ggplot(p8, aes(x=year, y=(`EV stock_BEV` + `EV stock_PHEV`) / 1e6, group = region, color = region)) +
  geom_line() + geom_point() + scale_x_continuous(breaks = seq(2010, 2022)) + 
  scale_y_continuous(
    labels = scales::label_number(suffix = " million"),
    breaks = seq(0, 30, by = 5)) +
  labs(
    title = "Electric car stocks, 2010 to 2022",
    x = "Year",
    y = "EV Stocks",
    color = "Region"
  )

world_map <- map_data("world") |> filter(region != 'Antarctica')
year_2022_data = wider_clean_df1 |> filter(year == 2022)
merged_data = merge(world_map, year_2022_data, by.x = 'region', 
                    by.y = 'region', all.x = T)
merged_data <- merged_data |>
  arrange(region, order)

ggplot(merged_data, aes(x = long, y = lat, group = interaction(region, group))) +
  geom_polygon(aes(fill = (`EV stock_BEV` + `EV stock_PHEV`) / 1e6), 
               color = NA, size = 0.1) + 
  geom_path(color = "black", size = 0.1) +
  scale_fill_gradient(low = "lightpink", high = "red3", 
                      na.value = "grey90", name = "EV Stocks (in millions)") +
  labs(title = "Electric car stocks, 2010 to 2022") +
  theme_void() + 
  coord_fixed() + theme(legend.position = "bottom") +
  guides(fill = guide_colourbar(title.position = "top", 
                                title.hjust = 0.5, 
                                barwidth = 20, 
                                barheight = 0.5, 
                                horizontal = TRUE)) 

```

This scatter plot and line graph shows the growth in the number of electric vehicles 
(in stock) across different regions from 2010 to 2022. China shows a notable increase 
here, with a steep climb starting around 2015 and reaching over 20 million vehicles 
by 2022, the highest out of the other regions. Out of these more highly populated 
regions, China seems to be an emerging leader in this space with the greatest stock 
of electric vehicles out of the other regions.


# Performing EDA and creating original visualizations using the data

# Diagram 15
```{r}

wider_clean_df1 |> 
  filter(!is.na(`Electricity demand_EV`)) |> 
  filter(!is.na(`Oil displacement Mbd_EV`)) |>
  ggplot() +
  geom_line(aes(x = year, y = `Electricity demand_EV`, color = "Electricity Demand")) +
  geom_line(aes(x = year, y = (`Oil displacement Mbd_EV`*100000), 
                color = "Oil Displacement")) +
  facet_wrap(~region, scales = "free_x") +
  theme_minimal() + 
  scale_x_continuous(breaks = seq(2010, 2022, by = 2)) +
  theme(axis.text.x = element_text(size = 6)) + 
  labs(title = "Electricity Demand and Oil Displacement, 2010 to 2022",
       x = "Year",
       y = "Electricity Demand (GWh) / Oil Displacement (Tens of Barrels per Day)") +
  theme(axis.title.y = element_text(size = 8)) +
  scale_color_manual(values = c("Electricity Demand"= "red", 
                                "Oil Displacement"= "blue")) +
  theme(legend.position = "bottom")

```
 
These faceted line graphs show how electricity demand and oil displacement changed between 
2010 and 2022 in China, Europe, India, USA, the World, and the rest of the World. These 
graphs show that as time progresses, both electricity demand and oil displacement increase in
all the regions, except India. However, there is only data on the oil displacement and 
electricity demand in India for 2019 through 2022, so it is hard to make any conclusions 
because there aren't enough years of data available to be able to effectively see if there 
have been any significant changes. This graph also shows that electricity demand and oil 
displacement tend to increase with each other, which may be because, as the world moves 
towards electric vehicles, demand for oil will be replaced by the demand for electricity.

# Diagram 16
```{r}

wider_clean_df1 |> 
  ggplot(aes(x= factor(year), y = `EV sales share_EV`)) +
  geom_boxplot(outlier.shape = NA) + 
  coord_flip() + 
  geom_point(aes(color = region), 
             position = position_jitter(width = 0.1, height = 0), 
             size = 0.75) +
  labs(title = "Distribution of Electric Vehicle Sales Share Per Year, 2010 to 2022",
       x = "Year",
       y = "EV Sales Share (%)",
       color = "Region") + 
  theme(legend.key.height = unit(5, 'mm'))

```

This boxplot shows the distribution of the percent of electric vehicle sales per year 
across all regions given in the dataset. This graph shows that as the years progress, 
the median electric vehicle sales share increases, which means that electric vehicles 
are taking up a larger percentage of vehicle sales. However, the range and interquartile 
range of electric vehicle sales share also increased with time, which means that there
is more variability in the sales share across regions. This illustrates that, as time 
passes, some regions have a lot of growth in electric vehicle sales share, while others 
do not, so the percentage of electric vehicle sales differs a lot more between regions 
in later years than in earlier years.

# Diagram 17
```{r}

df2_world_totals <- wider_clean_df2 |> 
  filter(region %in% "World") |>
  mutate(EV_total_sales = `EV sales_BEV` + `EV sales_PHEV`) |>
  mutate(total_sales = EV_total_sales / `EV sales share_EV` * 100) |>
  mutate(non_EV_total_sales = total_sales - EV_total_sales)

long_world2_totals <- df2_world_totals |>
  pivot_longer(cols = c(EV_total_sales, non_EV_total_sales), 
               names_to = "vehicle_type", 
               values_to = "sale_numbers")

ggplot(long_world2_totals, aes(x = year, y = sale_numbers, fill = vehicle_type)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("darkgreen", "lightyellow"),
                    labels = c("Electric cars", "Non-electric cars"),
                    name = "Vehicle Type") +
  labs(title = "Projected number of new cars sold, by type, World",
       x = "Year",
       y = "Total Cars Sold") +
  scale_y_continuous(labels = c("0 cars", "20 million cars", "40 million cars", 
                                "60 million cars", "80 million cars", "1 billion cars"),
                     breaks = seq(0, 10 * 10^7, by = 20 * 10^6)) +
  scale_x_continuous(breaks = unique(long_world2_totals$year)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```
  
This stacked bar graph shows the expected total number of cars sold in 2025 and 2030 
compared to the last couple of years in the data set. The graph reveals the trend that 
the amount of non-electric cars will likely stay the same, but the amount of electric
cars sold will go up. The graph shows that with a growing demand for vehicles, it is 
predicted that a larger percentage of people will go for electric cars rather than 
non-electric cars. 

# Diagram 18
```{r}

oil_regions_data <- wider_clean_df1 |>
  filter(region %in% c("India", "China", "Europe", "USA", 
                       "Rest of the world", "World")) |>
  select(region, year, `Oil displacement Mbd_EV`)

melted_data <- melt(oil_regions_data, 
                    id.vars = c("year", "region"), 
                    measure.vars = "Oil displacement Mbd_EV")

ggplot(melted_data, aes(x = region, y = year, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "brown1", high = "green") +
  labs(title = 
         "Oil Displacement(millions of barrels per day) Across Regions and Over Time", 
       x = "Region", y = "Year") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```
  
This heat map shows the oil displacement in millions of barrels per day for several 
regions over the years 2010 to 2022. The graph reveals that the world as a whole has
seen improvement in the amount of oil displacement, likely because of increased electric 
car use we've seen in other graphs. The graph also reveals that countries like India 
have not made much change at all and likely do not use electric vehicles very much. 

# Diagram 19
```{r}

wider_clean_df2 |> filter(region %in% c('USA', 'China', 'World', 'Europe')) |>
  ggplot(aes(x = year, 
         y = (`EV sales_BEV` + `EV sales_PHEV`) / 1e6, 
         size = `Oil displacement Mbd_EV`, 
         color = region)) +
  geom_point(alpha = 0.6) + scale_x_continuous(breaks = seq(2020, 2030)) +
  scale_y_continuous(breaks = seq(0, 40, by = 2)) + 
  scale_size_continuous(name = "Oil Displacement\n(Million barrels per day)") +
  labs(title = "Projected EV Sales and Oil Displacement Over Time", 
       x = "Year", y = "Number of EV's Sold (in millions)") +
  theme_minimal(base_size = 10) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8)
  )

```

Given the bubble time series scatter plot based on the projected data, we see that for 
highly populated regions such as China, Europe, and USA, the number of EV’s sold 
increases over time. For these regions as well as the world in of itself, as time 
progresses, the oil displacement also increases, suggesting a potential direct 
correlation between the number of EV’s sold and oil displacement. This correlation may 
suggest that more EV’s being sold can reduce oil consumption.

# Diagram 20
```{r}

ggplot(wider_clean_df2, aes(x = year, y = `EV sales share_EV`, fill = region)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~region, scales = "free_y") +
  labs(title = "Projected EV Sales Share by Region and Year",
       x = "Year",
       y = "EV Sales Share (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),  
        legend.position = "bottom") 

```

Given the bar chart for EV sales shares over time based on the projected data, we see that 
all of the listed countries share similar projections in EV sales shares. Specifically, 
we see an increase in EV sales shares across each of these countries by 2030, indicating 
that electric vehicles may have more of an overall presence within the market and road.
  
  
  
# Creating ML model for prediction

# 3 Diagram 21
```{r, warning=FALSE}

wider_clean_df1 |> select(`Electricity demand_EV`, `Oil displacement Mbd_EV`) |>
  filter(across(everything(), ~!is.na(.))) -> ml_filtered
ml_filtered = filter(ml_filtered, across(everything(), ~!is.na(.)))
mdl <- lm(`Oil displacement Mbd_EV` ~ `Electricity demand_EV`, data = ml_filtered)
a0 = coef(mdl)[1]
a1 = coef(mdl)[2]
ml_filtered = add_predictions(ml_filtered, mdl) 
ml_filtered
ggplot(ml_filtered, aes(x = `Electricity demand_EV`, `Oil displacement Mbd_EV`)) +
  geom_abline(aes(intercept = a0, slope = a1), color = "blue") +
  geom_point() + geom_point(aes(x = `Electricity demand_EV`, y = pred), color = "blue")
ml_filtered |> add_residuals(mdl) |> ggplot() + geom_histogram(aes(x=resid))
filter(wider_clean_df2, region == 'World', across(everything(), ~!is.na(.))) |>
  ggplot(aes(x = year, y = `Electricity demand_EV` / 1e3)) + 
  scale_x_continuous(breaks = seq(2020, 2030, by = 1)) +
  geom_point() + geom_smooth() + labs(y = "World Electricity Demand (in thousands)")

```

We narrowed down on electricity demand and oil displacement since we hypothesized that 
these two may share a correlation given how the increased use of electric vehicles 
implies a higher demand for electricity and increased oil displacement. We examined 
this initially through a scatter plot and noted that a reasonable linear model could 
be developed to predict future values. Given the charts, we observe that as the demand 
for electricity increases, we will likely see an increase in oil displacement (due to 
the increased presence of EV’s).
  
  
# Conclusions


The data set answers the questions of how quickly countries are moving towards electric cars,
and which countries are leading the way in reducing oil consumption. The graphs from the 
article show a positive trend toward more electric vehicles being sold throughout the world, 
with countries like Norway, Sweden, Germany, and China leading the way. The bar chart allows 
us to see that the majority of new electric cars sold tend to be fully battery electric, 
especially more recently, which is good news for carbon emissions. The number of new electric
cars sold has greatly increased in recent years, however, we see in the next graph that there
is still a lot of progress to be made, as electric cars are a very small fraction of the
total new cars sold. Cars currently in use is also a useful metric, as it can approximately
summarize car sales for up to the past decade. It takes time for the new sales of electric
cars to be reflected, so the numbers for current percentages of electric vehicles on the road
are lower. This can however give hope for lower emissions in years to come since we know that
the sales of electric cars have been rapidly increasing and so it is only a matter of time 
before that trend is reflected in current car use. Our first original graph shows the 
electricity demand and oil displacement and led us to the conclusion that they tend to 
increase with each other, which could be because as we are moving towards electric vehicles, 
the demand for oil will be replaced by the demand for electricity. We also looked at the 
distribution of electric vehicle sales per year across all the regions of the data set.  We 
looked at oil displacement as well for several regions including the world, and saw that the 
world saw improvement, but some regions had none at all like India. This tells us that some 
regions have made large increases in sales of electric vehicles, while others have made 
almost none at all, so there is a lot of variability in the responses of different regions to
the climate crisis. We also took a look at future data, which predicts no decrease in the 
amount of non-electric cars sold, but that electric car sales will continue to increase. We 
took a look at projected electric vehicles for several regions, and it is important
to note that the scales of the facet graphs are different so countries like India are still 
far behind China or the USA. We also sought to confirm the correlation between EVs being sold 
and oil displacement, and there seems to be a direct correlation suggesting that increasing 
EV sales can reduce oil consumption, something that was reinforced by our linear regression 
model. Overall, the world is taking positive steps towards reducing oil consumption, but 
there are some countries that are lagging behind in switching towards electric vehicles, 
which could be holding the rest of the world back from making even larger impacts on the 
environment. 


```{r}
library(tidyverse)
df1 <- read_csv('data/ev-historical.csv')
df2 <- read_csv('data/ev-steps-projected.csv')

```
